---
title: "Week Three"
subtitle: "Introduction to Tidyverse"
output: powerpoint_presentation
---

## Installing Tidyverse

```{r}
# install.packages("tidyverse")
```

## Loading Tidyverse

```{r}
library(tidyverse)
```

## Getting the data (base way)
```{r}
base_titanic <- read.csv("https://gist.githubusercontent.com/michhar/2dfd2de0d4f8727f873422c5d959fff5/raw/fa71405126017e6a37bea592440b4bee94bf7b9e/titanic.csv")
head(base_titanic)
```


## Getting the data (tidy way)

```{r cache=TRUE}
tidy_titanic <- read_csv("https://gist.githubusercontent.com/michhar/2dfd2de0d4f8727f873422c5d959fff5/raw/fa71405126017e6a37bea592440b4bee94bf7b9e/titanic.csv")
tidy_titanic
```

## Base class vs Tidy class
```{r}
class(base_titanic)
class(tidy_titanic)
```


## The filter verb
```{r}
filter(tidy_titanic, Age > 12)
```

## The filter verb
```{r}
tidy_titanic %>%
  filter(Age > 12)
```


## The select verb
```{r}
tidy_titanic %>%
  filter(Age > 12) %>%
  select(Survived, Sex)
```

## The group_by verb
```{r}
tidy_titanic %>%
  filter(Age > 12) %>%
  select(Survived, Sex) %>%
  group_by(Sex, Survived)
```



## The summarise verb
```{r}
tidy_titanic %>%
  filter(Age > 12) %>%
  select(Survived, Sex) %>%
  group_by(Sex, Survived) %>%
  summarise(num_sex_survive = n())
```

## The mutate verb
```{r}
tidy_titanic %>%
  filter(Age > 12) %>%
  select(Survived, Sex) %>%
  group_by(Sex, Survived) %>%
  summarise(num_sex_survive = n()) %>%
  group_by(Sex) %>%
  mutate(pct_sex_survive = num_sex_survive/sum(num_sex_survive) * 100)
```

## The mutate verb
```{r}
tidy_titanic %>%
  filter(Age > 12) %>%
  select(Survived, Sex) %>%
  group_by(Sex, Survived) %>%
  summarise(num_sex_survive = n()) %>%
  mutate(pct_sex_survive = num_sex_survive/sum(num_sex_survive) * 100)
```

## The mutate verb
```{r}
tidy_titanic %>%
  filter(Age > 12) %>%
  select(Survived, Sex) %>%
  group_by(Sex, Survived) %>%
  summarise(num_sex_survive = n()) %>%
  mutate(pct_sex_survive = num_sex_survive/sum(num_sex_survive) * 100) %>%
  filter(Survived == 1)
```


## Save in variable

```{r}
pct_surv <- tidy_titanic %>%
  filter(Age > 12) %>%
  select(Survived, Sex) %>%
  group_by(Sex, Survived) %>%
  summarise(num_sex_survive = n()) %>%
  mutate(pct_sex_survive = num_sex_survive/sum(num_sex_survive) * 100) %>%
  filter(Survived == 1)
pct_surv
```


## ggplot2 prt 1

```{r fig.height=5, fig.width=7, dpi=300}
ggplot(data = pct_surv)
```


## ggplot2 prt 2

```{r fig.height=5, fig.width=7, dpi=300}
ggplot(data = pct_surv, mapping = aes(x = Sex, y = num_sex_survive))
```

## ggplot2 prt 3

```{r fig.height=5, fig.width=7, dpi=300}
ggplot(data = pct_surv, mapping = aes(x = Sex, y = num_sex_survive)) + 
  geom_bar(stat = "identity")
```

## ggplot prt 4


```{r fig.height=5, fig.width=7, dpi=300}
ggplot(data = pct_surv, mapping = aes(x = Sex, y = pct_sex_survive)) + 
  geom_bar(stat = "identity")
```

## ggplot5

```{r fig.height=5, fig.width=7, dpi=300}
ggplot(data = pct_surv, mapping = aes(x = Sex, y = pct_sex_survive)) + 
  geom_bar(stat = "identity") +
  ylab("Percentage Surviving")
```



## ggplot6

```{r fig.height=5, fig.width=7, dpi=300}
ggplot(data = pct_surv, mapping = aes(x = Sex, y = pct_sex_survive)) + 
  geom_bar(stat = "identity") +
  ylab("Percentage Surviving") +
  labs(title = "Effect of Sex on Survival aboard the Titanic")
```



## ggplot7

```{r fig.height=5, fig.width=7, dpi=300}
ggplot(data = pct_surv, mapping = aes(x = Sex, y = pct_sex_survive)) + 
  geom_bar(stat = "identity") +
  ylab("Percentage Surviving") +
  labs(title = "Effect of Sex on Survival aboard the Titanic") +
  theme_classic()
```


## ggplot8

```{r fig.height=5, fig.width=7, dpi=300}
ggplot(data = pct_surv, mapping = aes(x = Sex, y = pct_sex_survive, fill = Sex)) + 
  geom_bar(stat = "identity") +
  ylab("Percentage Surviving") +
  labs(title = "Effect of Sex on Survival aboard the Titanic") +
  theme_classic() 
```


## ggplot9

```{r fig.height=5, fig.width=7, dpi=300}
ggplot(data = pct_surv, mapping = aes(x = Sex, y = pct_sex_survive, fill = Sex)) + 
  geom_bar(stat = "identity") +
  ylab("Percentage Surviving") +
  labs(title = "Effect of Sex on Survival aboard the Titanic") +
  theme_classic() +
  xlab(NULL)
```

## ggplot10

```{r fig.height=5, fig.width=7, dpi=300}
ggplot(data = pct_surv, mapping = aes(x = Sex, y = pct_sex_survive, fill = Sex)) + 
  geom_bar(stat = "identity") +
  ylab("Percentage Surviving") +
  labs(title = "Effect of Sex on Survival aboard the Titanic") +
  theme_classic() +
  xlab(NULL) +
  theme(legend.position = "none")
```


## ggplot11

```{r fig.height=5, fig.width=7, dpi=300}
ggplot(data = pct_surv, mapping = aes(x = Sex, y = pct_sex_survive, fill = Sex)) + 
  geom_bar(stat = "identity") +
  ylab("Percentage Surviving") +
  labs(title = "Effect of Sex on Survival aboard the Titanic") +
  theme_classic() +
  xlab(NULL) +
  theme(legend.position = "none") +
  scale_y_continuous(limits = c(0, 100), expand = c(0, 0))
```


## ggplot12

```{r fig.height=5, fig.width=7, dpi=300}
ggplot(data = pct_surv, mapping = aes(x = Sex, y = pct_sex_survive, fill = Sex)) + 
  geom_bar(stat = "identity") +
  ylab("Percentage Surviving") +
  labs(title = "Effect of Sex on Survival aboard the Titanic") +
  theme_classic() +
  xlab(NULL) +
  theme(legend.position = "none") +
  scale_y_continuous(limits = c(0, 100), expand = c(0, 0)) +
  scale_x_discrete(labels = c("Female", "Male"))
```


## ggplot13

```{r fig.height=5, fig.width=7, dpi=300}
ggplot(data = pct_surv, mapping = aes(x = Sex, y = pct_sex_survive, fill = Sex)) + 
  geom_bar(stat = "identity") +
  ylab("Percentage Surviving") +
  labs(title = "Effect of Sex on Survival aboard the Titanic") +
  theme_classic(base_size = 17) +
  xlab(NULL) +
  theme(legend.position = "none") +
  scale_y_continuous(limits = c(0, 100), expand = c(0, 0)) +
  scale_x_discrete(labels = c("Female", "Male"))
```


## ggplot14

```{r fig.height=5, fig.width=7, dpi=300}
my_figure <- ggplot(data = pct_surv, mapping = aes(x = Sex, y = pct_sex_survive, fill = Sex)) + 
  geom_bar(stat = "identity") +
  ylab("Percentage Surviving") +
  labs(title = "Effect of Sex on Survival aboard the Titanic") +
  theme_classic(base_size = 17) +
  xlab(NULL) +
  theme(legend.position = "none") +
  scale_y_continuous(limits = c(0, 100), expand = c(0, 0)) +
  scale_x_discrete(labels = c("Female", "Male"))
my_figure
```


## ggplot15

```{r fig.height=5, fig.width=7, dpi=300}
ggsave(plot = my_figure, filename = "my_fig.png", width = 7, height = 4.5)
```


## All together now

```{r fig.height=5, fig.width=7, dpi=300}
read_csv("https://gist.githubusercontent.com/michhar/2dfd2de0d4f8727f873422c5d959fff5/raw/fa71405126017e6a37bea592440b4bee94bf7b9e/titanic.csv") %>%  # Read the data into a tibble
  filter(Age > 12) %>%  # Only keep passengers with known Age
  select(Survived, Sex) %>%  # Only keep the Survived and Sex columns
  group_by(Sex, Survived) %>%  # Group by Sex and then Survived
  summarise(num_sex_survive = n()) %>%  # Count the number of passengers in these groups
  mutate(pct_sex_survive = num_sex_survive/sum(num_sex_survive) * 100) %>%  # Calculate the percent surviving
  filter(Survived == 1) %>%  # Only keep the surviving passengers
  ggplot(mapping = aes(x = Sex, y = pct_sex_survive, fill = Sex)) +  # Build the data and aesthetic layers 
  geom_bar(stat = "identity") +  # Build the geometric layer as a bar plot
  ylab("Percentage Surviving") +  # Make a Y axis label
  labs(title = "Effect of Sex on Survival aboard the Titanic") +  # Add a title
  theme_classic(base_size = 17) +  # Increase the font size and set a nice theme
  xlab(NULL) +  # Remove the X axis label
  theme(legend.position = "none") +  # Remove the legend
  scale_y_continuous(limits = c(0, 100), expand = c(0, 0)) +  # Set Y axis limits and remove the expansion
  scale_x_discrete(labels = c("Female", "Male")) +  # Capitalize the X axis Sex labels 
  ggsave(filename = "my_fig.png", width = 7, height = 4.5)
```



## Does Fare predict Survival?

```{r fig.height=5, fig.width=7, dpi=300}
read_csv("https://gist.githubusercontent.com/michhar/2dfd2de0d4f8727f873422c5d959fff5/raw/fa71405126017e6a37bea592440b4bee94bf7b9e/titanic.csv") %>%  # Read the data into a tibble
  ggplot(mapping = aes(x = Survived, y = Fare)) 
```


## Survived is being treated as a continuous variable

```{r fig.height=5, fig.width=7, dpi=300}
read_csv("https://gist.githubusercontent.com/michhar/2dfd2de0d4f8727f873422c5d959fff5/raw/fa71405126017e6a37bea592440b4bee94bf7b9e/titanic.csv") %>%  # Read the data into a tibble
  select(Survived)
```


## Factorizing Survived to make it discrete

```{r fig.height=5, fig.width=7, dpi=300}
read_csv("https://gist.githubusercontent.com/michhar/2dfd2de0d4f8727f873422c5d959fff5/raw/fa71405126017e6a37bea592440b4bee94bf7b9e/titanic.csv") %>%  # Read the data into a tibble
  ggplot(mapping = aes(x = factor(Survived), y = Fare)) 
```

## Add factor labels to improve readability

```{r fig.height=5, fig.width=7, dpi=300}
read_csv("https://gist.githubusercontent.com/michhar/2dfd2de0d4f8727f873422c5d959fff5/raw/fa71405126017e6a37bea592440b4bee94bf7b9e/titanic.csv") %>%  # Read the data into a tibble
  ggplot(mapping = aes(x = factor(Survived, labels = c("Died", "Survived")), y = Fare)) 
```


## Add the boxplot geometry

```{r fig.height=5, fig.width=7, dpi=300}
read_csv("https://gist.githubusercontent.com/michhar/2dfd2de0d4f8727f873422c5d959fff5/raw/fa71405126017e6a37bea592440b4bee94bf7b9e/titanic.csv") %>%  # Read the data into a tibble
  ggplot(mapping = aes(x = factor(Survived, labels = c("Died", "Survived")), y = Fare)) +
  geom_boxplot() 
```


## Coordinates: log-scale for y-axis

```{r fig.height=5, fig.width=7, dpi=300}
read_csv("https://gist.githubusercontent.com/michhar/2dfd2de0d4f8727f873422c5d959fff5/raw/fa71405126017e6a37bea592440b4bee94bf7b9e/titanic.csv") %>%  # Read the data into a tibble
  filter(Fare > 0) %>%
  ggplot(mapping = aes(x = factor(Survived, labels = c("Died", "Survived")), y = Fare)) +
  geom_boxplot() +
  scale_y_log10()
```


## Is age a confounding variable here? Make age brackets

```{r}
read_csv("https://gist.githubusercontent.com/michhar/2dfd2de0d4f8727f873422c5d959fff5/raw/fa71405126017e6a37bea592440b4bee94bf7b9e/titanic.csv") %>%  
  filter(is.numeric(Age) & Fare > 0) %>%
  mutate(Age_bracket = case_when(
    Age > 60 ~ "Senior",
    Age < 16 ~ "Child",
    TRUE ~ "Adult"  # This TRUE is equivalent to "else" in if...else
  ))
```

## Facet: split the plot by age bracket

```{r fig.height=5, fig.width=11, dpi=300}
read_csv("https://gist.githubusercontent.com/michhar/2dfd2de0d4f8727f873422c5d959fff5/raw/fa71405126017e6a37bea592440b4bee94bf7b9e/titanic.csv") %>%  # Read the data into a tibble
  filter(is.numeric(Age) & Fare > 0) %>%
  mutate(Age_bracket = case_when(
    Age > 60 ~ "Senior",
    Age < 16 ~ "Child",
    TRUE ~ "Adult"  # This TRUE is equivalent to "else" in if...else
  )) %>%
  ggplot(mapping = aes(x = factor(Survived, labels = c("Died", "Survived")), y = Fare)) +
  geom_boxplot() +
  scale_y_log10() +
  facet_wrap(~ Age_bracket) 
```


## Install and load ggpubr

```{r}
# install.packages("ggpubr")
library(ggpubr)
```


## Relationship between Age and Fare Price2


```{r fig.height=5, fig.width=11, dpi=300}
read_csv("https://gist.githubusercontent.com/michhar/2dfd2de0d4f8727f873422c5d959fff5/raw/fa71405126017e6a37bea592440b4bee94bf7b9e/titanic.csv") %>%  # Read the data into a tibble
  filter(is.numeric(Age) & Fare > 0) %>%
  mutate(Age_bracket = case_when(
    Age > 60 ~ "Senior",
    Age < 16 ~ "Child",
    TRUE ~ "Adult"  # This TRUE is equivalent to "else" in if...else
  )) %>%
  ggplot(mapping = aes(x = factor(Survived, labels = c("Died", "Survived")), y = Fare)) +
  geom_boxplot() +
  scale_y_log10() +
  facet_wrap(~ Age_bracket) +
  stat_compare_means(method = "t.test")
```


## Relationship between Age and Fare Price3


```{r fig.height=5, fig.width=11, dpi=300}
read_csv("https://gist.githubusercontent.com/michhar/2dfd2de0d4f8727f873422c5d959fff5/raw/fa71405126017e6a37bea592440b4bee94bf7b9e/titanic.csv") %>%  # Read the data into a tibble
  filter(is.numeric(Age) & Fare > 0) %>%
  mutate(Age_bracket = case_when(
    Age > 60 ~ "Senior",
    Age < 16 ~ "Child",
    TRUE ~ "Adult"  # This TRUE is equivalent to "else" in if...else
  )) %>%
  ggplot(mapping = aes(x = factor(Survived, labels = c("Died", "Survived")), y = Fare)) +
  geom_boxplot() +
  scale_y_log10() +
  facet_wrap(~ Age_bracket) +
  stat_compare_means(method = "t.test", label = "p.signif")
```


## Relationship between Age and Fare Price4


```{r fig.height=5, fig.width=11, dpi=300}
read_csv("https://gist.githubusercontent.com/michhar/2dfd2de0d4f8727f873422c5d959fff5/raw/fa71405126017e6a37bea592440b4bee94bf7b9e/titanic.csv") %>%  # Read the data into a tibble
  filter(is.numeric(Age) & Fare > 0) %>%
  mutate(Age_bracket = case_when(
    Age > 60 ~ "Senior",
    Age < 16 ~ "Child",
    TRUE ~ "Adult"  # This TRUE is equivalent to "else" in if...else
  )) %>%
  ggplot(mapping = aes(x = factor(Survived, labels = c("Died", "Survived")), y = Fare)) +
  geom_boxplot() +
  scale_y_log10() +
  facet_wrap(~ Age_bracket) +
  stat_compare_means(method = "t.test", label = "p.signif", comparisons = list(c("Died", "Survived")))
```



## Relationship between Age and Fare Price5


```{r fig.height=5, fig.width=11, dpi=300}
read_csv("https://gist.githubusercontent.com/michhar/2dfd2de0d4f8727f873422c5d959fff5/raw/fa71405126017e6a37bea592440b4bee94bf7b9e/titanic.csv") %>%  # Read the data into a tibble
  filter(is.numeric(Age) & Fare > 0) %>%
  mutate(Age_bracket = case_when(
    Age > 60 ~ "Senior",
    Age < 16 ~ "Child",
    TRUE ~ "Adult"  # This TRUE is equivalent to "else" in if...else
  )) %>%
  ggplot(mapping = aes(x = factor(Survived, labels = c("Died", "Survived")), y = Fare, fill =  factor(Survived, labels = c("Died", "Survived")))) +
  geom_boxplot() +
  scale_y_log10(limits = c(2, 1000)) +
  facet_wrap(~ Age_bracket) +
  stat_compare_means(method = "t.test", label = "p.signif", comparisons = list(c("Died", "Survived"))) +
  xlab(NULL) +
  ylab("Ticket Price (log scale)") +
  theme_bw(base_size = 17) +
  theme(legend.position = "none") +
  scale_fill_manual(values = c("goldenrod", "skyblue")) +
  ggsave(filename = "my_figure.png", height = 5, width = 11)
```

## Final example: correlation

```{r}
iris 
```


## Final example: correlation prt 2

```{r}
iris %>%
  filter(Species == "virginica") %>%
  ggplot(mapping = aes(x = Sepal.Length, y = Petal.Length)) +
  geom_point()
```


## Final example: correlation prt 2

```{r}
iris %>%
  filter(Species == "virginica") %>%
  ggplot(mapping = aes(x = Sepal.Length, y = Petal.Length)) +
  geom_point() +
  geom_smooth(method = "lm") 
```

## Final example: correlation prt 3

```{r}
iris %>%
  filter(Species == "virginica") %>%
  ggplot(mapping = aes(x = Sepal.Length, y = Petal.Length)) +
  geom_point() +
  geom_smooth(method = "lm") +
  stat_cor()
```


## Final example: correlation prt 4

```{r}
iris %>%
  ggplot(mapping = aes(x = Sepal.Length, y = Petal.Length, color = Species)) +
  geom_point() 
```


## Final example: correlation prt 5

```{r}
iris %>%
  ggplot(mapping = aes(x = Sepal.Length, y = Petal.Length, color = Species)) +
  geom_point() +
  geom_smooth(method = "lm") +
  stat_cor()
```


## Final example: correlation prt 6

```{r}
iris %>%
  ggplot(mapping = aes(x = Sepal.Length, y = Petal.Length, color = Species)) +
  geom_point() +
  geom_smooth(method = "lm") +
  stat_cor() +
  theme_classic(base_size = 16) +
  ylab("Petal Length (cm)") +
  xlab("Sepal Length (cm)") + 
  labs(title = "Petal vs Sepal Length in Iris Species") +
  ggsave("iris_lengths.png", height = 5, width = 8)
```




